Pre-processing Data from Embedded Questions NoContext Corpus Study

### 2021-11-11 analysis/06_experiment/rscripts/pre-processing.Rmd
added initial files from old repo

---
title: "Pre-processing Data from Embedded Questions Corpus Study"
author: Morgan Moyer
date: March 22, 2021
output: html_document
---

@@

# Read in data
```{r import data}
# Read database
corp = read.table("../../../corpus/results/swbd.tab",sep="\t",header=T,quote="")
# Read experimental data in
d = read.csv("../data/exp04_main-merged.csv")
nrow(d) # 174549
```

## Combine with database
```{r combine raw data with corpus file}
# rename item_id from corpus file
names(corp)[names(corp) == "Item_ID"] <- "tgrep_id"

@@

# join dfs together
d <- left_join(d, corp_match, by="tgrep_id")
# assign the negative values 0
d$rating[d$rating < 0] = 0
nrow(d) # 174549
```

```{r take a look at the raw distribution, include = FALSE}
ggplot(d, aes(x=rating)) +
  geom_histogram()

summary(d) # this doesn't show any negative rating values....
tail(d)
```

# Native speakers
## We exclude any speakers who did not report English as a native language in the post-survey demographic questions.
```{r Filter out non-native English speakers, warning = FALSE}
nrow(d) # 174549
length(unique(d$workerid)) # 1073
#filter out participants who did not report English as native language
d <- d %>%
  mutate(language = tolower(subject_information.language)) %>%
  filter(., grepl("eng", language))

# 25 participants removed for reporting non-english native language
nrow(d) #170068
length(unique(d$workerid)) # 1048
```

@@

# Remove subjects who failed 2/6 controls
```{r}
# for each control trial type, create a binary measure of whether the
# trial was passed
t = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="[_]") %>%
  group_by(workerid,paraphrase,trial) %>%
  filter(trial %in% c("movie", "book"), paraphrase %in% c("the")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
# nrow(t) #2028
a1 = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="_") %>%
  group_by(workerid,paraphrase,trial) %>%
  filter(trial %in% c("novels", "cookies"), paraphrase %in% c("all")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
# nrow(a1) #2028
a2 = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="_") %>%
  group_by(workerid,paraphrase,trial) %>% 
  filter(trial %in% c("tissue", "napkin"), paraphrase %in% c("a")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
# nrow(a2) #2028

@@

# How many removed using this criterion
length(unique(failed_controls$workerid)) # 37
length(unique(failed_controls$workerid))/length(unique(d$workerid))*100 # 3.5%

```
# Look at raw test items
```{r}
test = d %>%
  # Remove participants who fail 2/6 controls
  filter(!workerid %in% c(failed_controls$workerid)) %>%
  # Remove non-test items
  filter(!tgrep_id %in% c("example1", "example2", "example3", "example4","bot_check")) %>%
  filter(!grepl("control",tgrep_id)) %>%
  # Make a single matrix verb column
  unite("MatrixVerb", c("MatrixPredVerb", "MatrixPredOther", "MatrixPredParticle"), na.rm = TRUE, sep = " ", remove = FALSE) %>%
  # Remove leading and trailing whitespace
  mutate(MatrixVerb = str_trim(MatrixVerb))

@@

# Assign 0 to NaN
test[is.na(test)] <- 0
nrow(test) # 113632
```

## Recoding 
* 15 cases where MatrixVerbs weren't printed with TDTLite
* Modals
* QuestionType
```{r}
# 61188:30 --> funny 
test$MatrixVerb[test$tgrep_id == "61188:30"] = "funny"
test$VerbLemma[test$tgrep_id == "61188:30"] = "funny"
test$MatrixPredVerb[test$tgrep_id == "61188:30"] = "funny"

@@

# Recode nonfinite clauses as ModalPresent for graphs
test$ModalPresent[test$Finite == "no"] = "yes"
test$Modal[test$Finite == "no"] = "nonfinite"
# Recode contracted Modals
test$Modal[test$Modal == "ca"] = "can"
test$Modal[test$Modal == "'ll"] = "will"
test$Modal[test$Modal == "'d"] = "could"

test[is.na(test$EmbeddedSQ)] <- "no"
test$QuestionType = as.factor(test$QuestionType)

test$EmbeddedSQ = as.character(test$EmbeddedSQ)
test$QuestionType = as.character(test$QuestionType)
test$QuestionType[test$EmbeddedSQ == "yes"] <- "embeddedSAI"
test$EmbeddedSQ = as.factor(test$EmbeddedSQ)
test$QuestionType = as.factor(test$QuestionType)

table(test$QuestionType)
length(unique(test$workerid))
length(unique(test$workerid[test$QuestionType!="embeddedSAI"])) 
```

## Graph raw test items
```{r, eval=FALSE, include=FALSE}
agr = test %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position="dodge", show.legend = FALSE) +
  # ggtitle("Overall mean rating for each paraphrase") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/test_everything.pdf")

length(unique(test$MatrixVerb)) # 175 verbs

@@

## Remove rhetorical questions
```{r}
test_agr = test %>%
  group_by(tgrep_id, paraphrase) %>%
  summarize(mean_rating = mean(rating))

other_ratings = test %>%
  group_by(tgrep_id, Sentence, paraphrase) %>%
  summarize(mean_rating = mean(rating)) %>%
  filter((mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="a"]) & 
           (mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="every"]) & 
           (mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="the"]))

View(other_ratings)

# How much data removed
nrow(other_ratings)/nrow(test_agr)*100 # 15.5%
nrow(test_agr) # 4052

or_ids = other_ratings$tgrep_id
test_other = test %>%
  filter(tgrep_id %in% or_ids)
nrow(test_other)/nrow(test)*100 # 15.5

test_other %>%
  filter(paraphrase == "other") %>%
  group_by(Sentence, paraphrase) %>%
  summarize(mean_rating = mean(rating))  # %>%
  View()

@@

nrow(test_norm)/nrow(test)*100 # 84%

nrow(test_norm) # 95956

# write.csv(test_norm,"../data/embedded_no_rhetorical.csv")
```

# Looing at QuestionType
## Look at SAI and Particle Verbs in Rhetorical Qs
```{r, include = FALSE}
length(unique(test_other$Sentence)) # 159 items

# Rhetorical questions with embedded SAI
test_removed_sq = test_other %>%
  filter(EmbeddedSQ == "yes")
length(unique(test_removed_sq$Sentence)) # 12 items
nrow(test_removed_sq)/nrow(test_other)*100 # 7.5%

ggplot(test_removed_sq, aes(x=VerbLemma)) +
  geom_histogram(stat="count") +
  theme(axis.text.x = element_text(angle = 90))

# Rhetorical questions with verbs particles
test_removed_pp = test_other %>%
  filter(MatrixPredParticle != "")
length(unique(test_removed_pp$Sentence)) # 29 items
nrow(test_removed_pp)/nrow(test_other)*100 #18%
```

## Look at SAI and Particle Verbs in Test Qs
```{r, include = FALSE}
# In the test data, look at embedded SAI
sq = test_norm %>%
  filter(EmbeddedSQ == "yes")

length(unique(sq$Sentence)) # 58 items
nrow(sq)/nrow(test_norm)*100 # 6.9%

# In the test data, look at verbs with Particles
pp = test_norm %>%
  filter(MatrixPredParticle != "")
length(unique(pp$Sentence)) # 49 items
nrow(pp)/nrow(test_norm)*100 # 5.7%

# Historgram of verbs
ggplot(sq, aes(x=VerbLemma)) +
  geom_histogram(stat="count") +
  theme(axis.text.x = element_text(angle = 90))

View(unique(pp[,c("tgrep_id","Sentence","VerbLemma","Question")]))
View(unique(test_norm[,c("tgrep_id","Sentence","Question")]))
table(test_norm$EmbeddedSQ, test_norm$Wh)
```

## Look at questions in test data without embedded SAI

```{r, include = FALSE}
# In the test data questions without embedded SAI
no_sq = test_norm %>%
  filter(EmbeddedSQ != "yes")

nrow(no_sq)/nrow(test_norm)*100 #93%
View(unique(no_sq[,c("tgrep_id","Sentence","Question")]))

no_pp = test_norm %>%
  filter(MatrixPredParticle == "")
nrow(no_pp)/nrow(test_norm)*100 #94%

# take a look individual questions based on matrix verb frequency
t = no_sq %>%
  select(tgrep_id,VerbLemma,Sentence, Question) %>%
  group_by(VerbLemma) %>%
  mutate(count = n()) %>%
  filter(VerbLemma == "know") %>%
  unique() %>%
  View()

test_norm %>%
  filter(tgrep_id == "173250:33") %>%
  group_by(Sentence,Question,paraphrase) %>%
  summarize(mean_rating = mean(rating), sd = sd(rating)) %>%
  View()

# CASES THAT SEEM LIKE NOT REAL EMBEDDINGS
# 171379:158 --> based on what they see *t*-5
# but people who *t*-4 d-, don't have day in and day out life long experience with what prison is really like *t*-1 and what the options are *t*-2, * to ask them *-3 to decide for one case, on one person, based on what they see *t*-5, i think 0 that is maybe, uh, a little bit naive.

# 31019:97 --> reshot where they could
# and they just change the title and, uh, and reshot a few parts, you know, where they, yeah, you know, where they could *?* *t*-1.

# 
# XXXX
# 37675:66 coming to a point soon when....
# i, you know, i think that there's coming to a point *ich*-1 real soon when ticket prices are going *-2 to be to the point where the average fan can't go *t*-3 *t*-4
# 
# 177034:133 --> witnessed what we did
# 
# uh, and th-, and the problems that, that *t*-3 has come that *t*-4 has come from this decision that that jury came to *t*-1 after all of us witnessed what we did *t*-2 with the video tape beating. uh just makes you wonder i guess whether or not they,
# 
# XXXX
# 56988:40 
# in fact, it *exp*-2 was even pretty much spelled *-1 out who you did vote for *t*-3, up until fairly recently
# 
# XXXXXX
# 57547:40 
# i tried *-1 to remember *ich*-3 as i took stuff off where it went *t*-2,
# 
# ???
# 80505:155 --> precise about how you do it 
#   because if you, uh, use drop cloths *-5 to cover everything and you use masking tape *-6 to trim out, uh, all the parts that you don't want *-1 to slop over onto *t*-2, you don't have *-3 to be quite as precise about how you do it *t*-4
# 
#   XXXX
#   160971:117
#   i mean, i just read something that, two thirds or three quarters, i'm not exactly sure of our national debt are, was created *-1 during the reagan bush *ich*-3 era who *t*-2 are suppo-, supposedly fiscally conservative.

the_high = no_sq %>%
  filter((paraphrase == "the")) %>% #  & (Wh == "why")
  group_by(tgrep_id,Sentence,VerbLemma,Question) %>%
  summarize(mean_rating = mean(rating), sd = sd(rating)) %>%
  filter(mean_rating > .5) %>%
  View()

# 94825:83
# and when he started, uh, at skyline high school t-2, i asked him, son, when, uh, you know, where do i go -1 to sign up for the p t a t-3?
# 	
# 56723:24
# im trying 1 to think what they call t-2 their soup.

a_high = no_sq %>%
  filter((paraphrase == "a") & (VerbLemma == "know")) %>% 
  group_by(tgrep_id,Sentence,VerbLemma,Question) %>%
  summarize(mean_rating = mean(rating), sd = sd(rating)) %>%
  filter(mean_rating > .3) %>%
  View()

# 29177:41
# and, uh, i talked to a thirteen year old *ich*-2 last night who *t*-1's, uh, goes to, uh, one of the plano high schools or junior highs, i guess,
# 	
# 34493:13
# so, if and when we ever get a baby-sitter *t*-1, like we did with silence of the lambs, we go *-2 see it.
# 	
# 109916:100
# and, uh, i just, it *exp*-4's just weird that i was just talking to somebody *ich*-3 this morning who *t*-1's, who *t*-2's a working, uh, lady, a working mom,
# 
# 90258:41
# she has a, a child *ich*-2 in connecticut who *t*-1 is extremely gifted,
# 16096:37
# well, it says, why would you, what you would have *t*-1 for a dinner party.
# 85372:42
# um, there are some areas *ich*-2 though where, where they have a problem with, uh, the houses kind of sinking a little bit into the, to the, uh *t*-1,
# 170922:29
# and the money *ich*-3 is also another issue. how you're going *-1 to pay for it *t*-2.
# 83586:41
# and then i gave it to a guy *ich*-3 *-4 to repair who *t*-1 works for the aut-, for the audi dealer
# 79010:88
# but recently, uh, the refuse department has, has, uh, is working on *-1 getting a system *ich*-3 going where we will actually have, uh, a bend that we put things in *t*-2 for * recycling these other things *t*-4
# ??31019:97
# and they just change the title and, uh, and reshot a few parts, you know, where they, yeah, you know, where they could *?* *t*-1.

all_high = no_sq %>%
  filter(paraphrase %in% c("every")) %>% #  & (Wh == "when") & (ModalPresent == "yes")
  group_by(tgrep_id,Sentence,VerbLemma,Question) %>%
  summarize(mean_rating = mean(rating), sd = sd(rating)) %>%
  filter(mean_rating > .5) %>%
  View()

# 66692:37 --> careful
# i have got *-1 to be so much more careful with what i do *t*-2.
# 
# 61922:110
# and and it puts me right up-to-date what i'm, what i have *-1 to do *t*-2, what i have *-3 to ship *t*-4, what i should be expecting *t*-5 to be returned *-7.

other_high = no_sq %>%
  filter(paraphrase %in% c("other")) %>%
  group_by(tgrep_id,Sentence,VerbLemma,Question) %>%
  summarize(mean_rating = mean(rating), sd = sd(rating)) %>%
  # filter(mean_rating > .5) %>%
  View()

part = no_sq %>%
  filter(MatrixPredParticle != "")
# What % of non-SAI Qs still have Particles?
nrow(part)/nrow(no_sq)*100 # 5%

no_pp_no_SAI = test_norm %>%
  filter((MatrixPredParticle == "") & (EmbeddedSQ != "yes"))
# What % of all test Qs are NOT SAI OR Particle?
nrow(no_pp_no_SAI)/nrow(test_norm)*100 # 88%

View(unique(no_pp_no_SAI[,c("VerbLemma","Question","Sentence")]))

## Graph Embedded Subject-Aux Inversion
```{r, include = FALSE}
agr = test_norm %>%
  group_by(EmbeddedSQ,paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, alpha=EmbeddedSQ, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=position_dodge(0.9)) +
  # facet_grid(EmbeddedSQ~Wh) +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  guides(alpha=guide_legend(title="Embedded SAI")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette) +
    scale_alpha_discrete(range = c(.5,1))
ggsave("../graphs/overall_SAI.pdf",width=3,height=3)
# ggsave("../graphs/modxwh_SAI.pdf",width=8,height=3)

agr = sq %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

dodge = position_dodge(.9)
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  ggtitle("overall ratings for SAI in test qs") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  ylim(0,.6) +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)

agr = sq %>%
  group_by(Wh,ModalPresent,paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, alpha=ModalPresent, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=position_dodge(0.9)) +
  facet_wrap(~Wh, ncol=2) +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  ggtitle("overall ratings for SAI in test qs") +
  guides(fill=FALSE) +
  guides(alpha=guide_legend(title="Modal present")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette) +
    scale_alpha_discrete(range = c(.5,1))

test_norm_noSQ = test_norm %>%
  filter(EmbeddedSQ != "yes") %>%
  group_by(VerbLemma, tgrep_id,Sentence,paraphrase) %>%
  summarize(count = n(),mean_rating = mean(rating))
View(test_norm_noSQ)

# Look at Matrix Verbs
```{r, eval=FALSE, include=FALSE}
verb_count = test %>%
  group_by(VerbLemma) %>%
  summarize(count = n())
View(verb_count)

@@

# Focusing just on "Every", "a" and "the"
## Normalize ratings to make a probability distribution

```{r }
# remove 'other' ratings
# critical with embeddedSQ = 71967
# critical withOUT embeddedSQ = 66963
critical = test_norm %>%
  filter(paraphrase %in% c("every","a","the"))
  # filter((MatrixPredParticle == "") & (EmbeddedSQ != "yes"))
# nrow(critical)
# View(critical)
# unique by-item/by-participant combo
critical$ids = paste(critical$workerid,critical$tgrep_id)

@@

# FIND OUT:
# by-participant variability using KL divergence from the mean

nrow(normed) # 64911
names(normed)

@@

# save to .csv to load into analysis script
# write.csv(normed,"../data/normed.csv")

## Graph normed data Overall
```{r}
agr = normed %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

dodge = position_dodge(.9)
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  ylim(0,.6) +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
ggsave("../graphs/overall_noSQ.pdf",width=2,height=2)
```

## Graph Interaction between Modal and Wh
```{r}

@@

# ggsave("../graphs/modxwh_noSQ.pdf",width=3,height=4)
```

## Look at just Wh
```{r}
agr = normed %>%
  group_by(paraphrase,Wh) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
# Re-Order the WH-levels by overall composition of DB
agr$Wh <- factor(agr$Wh, levels=c("what","how","where","why","who","when"))
ggplot(agr,aes(x=Wh, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9))  +
  # ggtitle("Mean rating for Wh-Word") +
  xlab("Wh-Word") +
  ylab("Mean rating") +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/final_normed_wh.pdf")
```

## Look at just ModalPresent
```{r}
agr = normed %>%
  group_by(paraphrase,ModalPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=ModalPresent, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9)) +
  xlab("Modal Present") +
  ylab("Mean rating") +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# legend.spacing.y = unit(-10, 'cm'))
# guides(fill=guide_legend(title="Paraphrase"))
# ggsave("../graphs/final_normed_modalpresent.pdf")

@@

## Graph MatrixVerb only
```{r}
verb_count = normed %>%
  group_by(VerbLemma) %>%
  summarize(count = n())

@@

surp = subset(normed,normed$VerbLemma == "surprise")
nrow(surp)
agr = normed %>%
  # Most frequent
  filter(VerbLemma %in% c("know","see","wonder","understand","be","understand","surprise","tell","figure","say","think","sure","learn","remember")) %>%
  group_by(paraphrase,VerbLemma,ModalPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

@@

```{r Looking at individual items}
the_high = normed %>%
  filter((paraphrase == "the")) %>% #  & (Wh == "why")
  group_by(tgrep_id,Sentence) %>%
  summarize(mean_rating = mean(normed_rating), sd = sd(normed_rating)) %>%
  filter(mean_rating > .5) %>%
  View()

>>>>

---
title: "Pre-processing Data from Embedded Questions NoContext Corpus Study"
author: Morgan Moyer
date: May 18, 2021
output: html_document
---

@@

```{r import data}
# Read database
corp = read.table("../../../corpus/results/swbd.tab",sep="\t",header=T,quote="")
# Read experimental data in
d = read.csv("../../05_experiment/data/exp0506-merged.csv")
d <- d %>%
  filter(.,grepl("experiment_06",proliferate.condition))
nrow(d) # 108995
```

```{r combine raw data with corpus file,warning=FALSE}
# rename item_id from corpus file
names(corp)[names(corp) == "Item_ID"] <- "tgrep_id"

@@

# join dfs together
d <- left_join(d, corp_match, by="tgrep_id")
# assign the negative values 0
d$rating[d$rating < 0] = 0
# table(d$proliferate.condition)
```

```{r take a look at the raw distribution, include=FALSE, echo=FALSE}
ggplot(d, aes(x=rating)) +
  geom_histogram()

# Native speakers
## We exclude any speakers who did not report English as a native language in the post-survey demographic questions.
```{r Filter out non-native English speakers, warning = FALSE}
nrow(d) # 104686
length(unique(d$workerid)) # 717
#filter out participants who did not report English as native language
d <- d %>%
  mutate(language = tolower(subject_information.language)) %>%
  filter(., grepl("eng", language))

# 28 participants removed for reporting non-English native language
nrow(d) #104686
length(unique(d$workerid)) # 689
```

@@

# Remove Subjects who failed 2/6 Controls
```{r Remove subjects who failed 2/6 controls}
# for each control trial type, create a binary measure of whether the
# trial was passed
t = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="[_]") %>%
  group_by(workerid,paraphrase,trial) %>%
  filter(trial %in% c("movie", "book"), paraphrase %in% c("the")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
# nrow(t) #2028
a1 = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="_") %>%
  group_by(workerid,paraphrase,trial) %>%
  filter(trial %in% c("novels", "cookies"), paraphrase %in% c("all")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
# nrow(a1) #2028
a2 = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="_") %>%
  group_by(workerid,paraphrase,trial) %>% 
  filter(trial %in% c("tissue", "napkin"), paraphrase %in% c("a")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
# nrow(a2) #2028

@@

# How many removed using this criterion
length(unique(failed_controls$workerid)) # 43
length(unique(failed_controls$workerid))/length(unique(d$workerid))*100 # 6.2%

# Look at raw test items

```{r Look at raw test items}
test = d %>%
  # Remove participants who fail 2/6 controls
  filter(!workerid %in% c(failed_controls$workerid)) %>%
  # Remove non-test items
  filter(!tgrep_id %in% c("example1", "example2", "example3", "example4","bot_check")) %>%
  filter(!grepl("control",tgrep_id)) %>%
  # Make a single matrix verb column
  unite("MatrixVerb", c("MatrixPredVerb", "MatrixPredOther", "MatrixPredParticle"), na.rm = TRUE, sep = " ", remove = FALSE) %>%
  # Remove leading and trailing whitespace
  mutate(MatrixVerb = str_trim(MatrixVerb))

@@

# Assign 0 to NaN
test[is.na(test)] <- 0
nrow(test) # 72216
```

## Recode 
* 15 cases where MatrixVerbs weren't printed with TDTLite
* Modals
```{r, Recode}
# 61188:30 --> funny 
test$MatrixVerb[test$tgrep_id == "61188:30"] = "funny"
test$VerbLemma[test$tgrep_id == "61188:30"] = "funny"
test$MatrixPredVerb[test$tgrep_id == "61188:30"] = "funny"

@@

# Recode nonfinite clauses as ModalPresent for graphs
test$ModalPresent[test$Finite == "no"] = "yes"
test$Modal[test$Finite == "no"] = "nonfinite"
# Recode contracted Modals
test$Modal[test$Modal == "ca"] = "can"
test$Modal[test$Modal == "'ll"] = "will"
test$Modal[test$Modal == "'d"] = "could"
```

# Graph Raw test items
```{r Graph raw test items, eval=FALSE, include=FALSE}
agr = test %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position="dodge", show.legend = FALSE) +
  # ggtitle("Overall mean rating for each paraphrase") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
ggsave("../graphs/test_everything.pdf")

length(unique(test$MatrixVerb)) # 193 verbs

@@

# Remove Rhetorical Questions

```{r Remove rhetorical questions}
test_agr = test %>%
  group_by(tgrep_id, paraphrase) %>%
  summarize(mean_rating = mean(rating))

other_ratings = test %>%
  group_by(tgrep_id, paraphrase) %>%
  summarize(mean_rating = mean(rating)) %>%
  filter((mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="a"]) & 
           (mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="every"]) & 
           (mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="the"]))

# How much data removed
nrow(other_ratings)/nrow(test_agr)*100 # 17.8%
nrow(test_agr) # 4772

or_ids = other_ratings$tgrep_id
test_other = test %>%
  filter(tgrep_id %in% or_ids)
nrow(test_other)/nrow(test)*100 # 18.6%

test_other %>%
  group_by(Sentence,Question,paraphrase,rating) %>%
  summarize(mean_rating = mean(rating)) %>%
  filter(paraphrase == "other") %>%
  View()

@@

nrow(test_norm)/nrow(test)*100 # 81%

nrow(test_norm) # 58756

# write.csv(test_norm,"../data/no_rhetorical_nocontext_embedded.csv")
```

```{r Look at Matrix Verbs, eval=FALSE, include=FALSE}
verb_count = test %>%
  group_by(VerbLemma) %>%
  summarize(count = n())
View(verb_count)

@@

# Normalize ratings

```{r Normalize ratings to make a probability distribution}
# remove 'other' ratings
critical = test_norm %>%
  filter(paraphrase %in% c("every","a","the"))

# unique by-item/by-participant combo
critical$ids = paste(critical$workerid,critical$tgrep_id)

@@

@@

# save to .csv to load into analysis script
# write.csv(normed,"../data/normed_nocontext_embedded.csv")
```

# Graph normed data

## Overall

```{r Graph normed data Overall}
agr = normed %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

dodge = position_dodge(.9)
ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position=dodge,stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=dodge, show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  ylim(0,.6) +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/overall_nocontext_embedded.pdf",width=2,height=2)
```

## Modal x Wh-Word

```{r Look at Interaction between Modal and Wh}

@@

# ggsave("../graphs/modxwh_nocontext_embedded.pdf",width=3,height=4)
```

```{r Look at just Wh,include=FALSE}
agr = normed %>%
  group_by(paraphrase,Wh) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
# Re-Order the WH-levels by overall composition of DB
agr$Wh <- factor(agr$Wh, levels=c("what","how","where","why","who","when"))
ggplot(agr,aes(x=Wh, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9))  +
  # ggtitle("Mean rating for Wh-Word") +
  xlab("Wh-Word") +
  ylab("Mean rating") +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
```

```{r Look at just ModalPresent,include=FALSE}
agr = normed %>%
  group_by(paraphrase,ModalPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=ModalPresent, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9)) +
  xlab("Modal Present") +
  ylab("Mean rating") +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)

@@

```{r Graph MatrixVerb only}
verb_count = normed %>%
  filter(grepl("experiment_06",proliferate.condition)) %>%
  group_by(VerbLemma) %>%
  summarize(count = n())

@@

agr = normed %>%
  # Most frequent
  filter(VerbLemma %in% c("know","see","wonder","understand","be","understand","surprise","tell","figure","say","think","sure","learn","remember")) %>%
  group_by(paraphrase,VerbLemma,ModalPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

@@

```{r Looking at individual items}
the_high = normed %>%
  filter((paraphrase == "the") & (VerbLemma == "surprise")) %>% #  & (Wh == "why")
  group_by(tgrep_id,Sentence) %>%
  summarize(mean_rating = mean(normed_rating), sd = sd(normed_rating)) %>%
  filter(mean_rating > .5) %>%
  View()
